### 2022-08-08 article.tex
Improve

@@

There are multiple ways of publishing code,
that are increasingly hard to do, yet result
in more reproducible experiments:
(1) publishing the code as text only, e.g. by printing
it in a Supplementary Materials, (2) publishing
the code on a hosted website, e.g on GitHub, with
added features (3) publishing 
the installed code within a virtual environment,
for example, using a Singularity container.
Here I will go into details of each of these ways.

@@

A feature that deserves more attention
is that code can actively teach what it does.
A mechanism to do so, which is present in most programming
languages, is error handling: errors are the mechanism
by which a programmer can teach the user what to do/expect.
One example is when calculating the variance of a distribution.
To calculate the variance of a distribution, at least two values
are required. When the variance of only one value is requested,
an error message can helpfully indicate that one needs at least two values.

@@

Publishing raw code is already better than not publishing code at all,
as that raw code can be assumed to have run successfully at least once.
To reproduce an experiment from raw code can be painful, 
yet way better than rewriting the same code from scratch.

@@

Code is sometimes published alongside a paper, 
in the form of a hosted website.
This way of publishing code gives additional options,
both for the authors of that code, as well as its users.

@@

There are multiple websites for hosting code, with BitBucket, GitHub,
GitLab and SourceForge being the most well-known.
Of these, GitHub is the most popular website to host code in general,
with millions of users.
GitHub users can create dedicated websites (called 'repositories')
to upload code. GitHub hosts that uploaded by and provides
a user interface for every visiter of such a website.
The use of code hosting websites, such as GitHub,
accommodates collaboration (\cite{perez2016ten}).
\begin{sloppypar}\url{https://github.com/richelbilderbeek/babette},\end{sloppypar}
and improves transparency (\cite{gorgolewski2016practical}).
An example from the author is the website \cite{bbbqarticleissue157},
that hosts part of the code for the paper \cite{bilderbeek2022transmembrane},
as shown in Figure \ref{fig:bbbqarticleissue157}.

@@

Hosted code has many additional features, such as
a version-controlled history, continuous integration scripts
and automatically generated metrics.

@@

Hosted code commonly has a commit history. For example, GitHub
uses a version control system called \verb|git| (hence its name).
In practice, this means that when a change is made to the code,
a new version is created. In the case that the change was harmful,
one can go back to an earlier version and continue again from there.

@@

Hosted code improves transparency (\cite{gorgolewski2016practical}).
The history of changes, called the commit history, 
helps to assess the honesty in the reporting of the code. 
Ideally, the full version history of the code 
has been generated, by hosting the code directly at its inception.
However, in some repositories, only the final version of the code has been
posted, which means all historical information is lost.
From a code history, it can be observed, for example, 
how much statistical tests have really/actually been done 
and see if all are reported:
not reporting a statistical tests is considered a Dubious Research Practice 
\richel{REF}, as this allows dishonest reporting of p-value hacking.

>>>>

One pseudorandom example of a mismatch between code and article
was found by me in \cite{labrecque2019interpretation}.
After the corresponding author of that paper was contacted,
the code was updated and confirmed to have the same
results as published in the paper.

@@

As it is the code that is the most important actor in an the experiment,
it should be treated as equally or more important than the scientific
paper describing what is does.
Regardless of this, most academic journals, 
when concerning computational experiments
only review the (usually) English description of the code
and make the review of code optional.
One notable exception is rOpenSci (\cite{ram2018community}),
which is an organisation that does peer-review academic code 
and facilitates the optional publication of a paper.
Although both code and paper are peer-reviewed,
most effort is spent on reviewing the code,
as can be seen in, for example, 
the \verb|mcbette| R package (\cite{bilderbeek2020mcbette}).

There are multiple ways of publishing code,
that require increasingly more knowledge to do and result
in increasingly more reproducible experiments:
(1) publishing the code as text only, e.g. by printing
it in a Supplementary Materials, (2) publishing
the code on a hosted website, e.g on GitHub, with
added features (3) publishing 
the installed code within a virtual environment,
for example, using a Singularity container.
Here I will go into details of each of these ways.

@@

A feature that deserves more attention
is that 'code must act as a teacher for future developers' \cite{sadowski2018modern}.
Error handling is one of the mechanisms to do so.
For example, let is when calculating the variance of a distribution.
To calculate the variance of a distribution, at least two values
are required. When the variance of only one value is requested,
an error message can helpfully indicate that one needs at least two values.

@@

Publishing raw code is already better than not publishing code at all,
as that raw code can be assumed to have run successfully at least once.
To reproduce an experiment from raw code can be painful, 
yet it is usually better than rewriting the same code from scratch.

@@

Code is sometimes published alongside a paper, 
in the form of a hosted website.
This way of publishing code gives additional options,
both for the authors of that code, as well as its users.
This practice has increased,
with the abstracts from articles in the academic journal
'Bioinformatics' referring to such a hosted website
has increased tenfold, going from below 5 percent in 2009 
to around 50 percent in 2017 (\cite{russell2018large}).

@@

There are multiple websites for hosting code, with BitBucket, GitHub,
GitLab and SourceForge being the most well-known.
Of these, GitHub is the most popular website to host code in general,
with millions of users.
GitHub users can create dedicated websites (called 'repositories')
to upload code. Code hosts, such as GitHub, 
store that uploaded code and provide ways 
for visiters to download the code and/or interact with it.
The use of such code hosting websites
accommodates collaboration (\cite{perez2016ten}).
and improves transparency (\cite{gorgolewski2016practical}).
An example from the author is the website \cite{bbbqarticleissue157},
that hosts part of the code for the paper \cite{bilderbeek2022transmembrane},
as shown in Figure \ref{fig:bbbqarticleissue157}.

@@

Hosted code has many additional features, such as
a version-controlled history, continuous integration scripts
and automatically generated metrics.
A code hosting website produces and shares this metadata.
This metadata can help to determine the
reproducibility, correctness and quality of code
of computational experiments.

@@

Hosted code commonly has a commit history. For example, GitHub
uses a version control system called \verb|git| (hence its name).
In practice, this means that when a change is made to the code,
a new version is created. In the case that the change was harmful,
one can go back to an earlier version and continue again from there.
It is a general recommendation to put version control
on all human-produced data (\cite{wilson2014best}).

@@

Hosted code improves transparency (\cite{gorgolewski2016practical}).
The history of changes, called the commit history, 
helps to assess the honesty in the reporting of the code. 
Ideally, the full version history of the code 
has been generated, by hosting the code directly at its inception.
However, in some repositories, only the final version of the code has been
posted, which means all historical information is lost.
From a code history, it can be observed, for example, 
how much statistical tests have really/actually been done 
and see if all are reported:
failing to reporting a (usually non-significant) statistical tests is 
one of the many degrees of freedom 
in p-value hacking (\cite{wicherts2016degrees}).
